---
title: "Count regression"
author: "Vera Miler Jerkovic"
date: "August 4, 2019"
output:
  pdf_document: default
  html_document: default
  rmarkdown::github_document
---

rm(list=ls())
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(kableExtra)
library(grid)
library(gridExtra)
library(knitr)
library(magrittr)
library(tidyr)
library(ggcorrplot)
library(psych)
library(car)
library(MASS)
library(VIM)
library(DataExplorer)
library(reshape2)
library(readODS)
library(e1071)
library(caret)
library(foreign)
library(lubridate)
```

```{r,echo=TRUE}
df<-read_ods('C:\\Users\\Vera\\Desktop\\nez-opendata-20193-20190325.ods', na="")
df<-df[,-7]
colnames(df)<-c("ID", "Date and Time", "WGS_X", "WGS_Y", "Outcome", "description")
```

```{r,echo=TRUE}
summary(df)
str(df)
unique(df$Outcome)
unique(df$description)
#create_report(baza)
```

#Separate Date and Time
```{r,echo=TRUE}
df <- separate(data = df, col = "Date and Time",
                    into = c('Date', 'Time'),
                    sep = ',',
                    remove = F)
df <- separate(data = df,
                    col = "Date",
                    into = c('Day', 'Month', 'Year'),
                    sep = '\\.',
                    remove = F)
df <- separate(data = df,
                    col = "Time",
                    into = c('Hour', 'Minute'),
                    sep = ':',
                    remove = F)

unique(df$Year)
df$Month[df$Month=="01"]<-"January"
df$Month[df$Month=="02"]<-"February"
```

#Belgrade is on: Coordinates: 44°49'N 20°28'E
#keep only data around those values
```{r,echo=TRUE}
range(df$WGS_X, na.rm = TRUE)
range(df$WGS_Y, na.rm = TRUE)
df_new <- df %>% 
  filter (WGS_X < mean(WGS_X, na.rm = TRUE) + 3*sd(WGS_X, na.rm = TRUE), 
          WGS_X > mean(WGS_X, na.rm = TRUE) -3*sd(WGS_X, na.rm = TRUE), 
          WGS_Y < mean(WGS_Y, na.rm = TRUE) + 3*sd(WGS_Y, na.rm = TRUE), 
          WGS_Y > mean(WGS_Y, na.rm = TRUE) - 3*sd(WGS_Y, na.rm = TRUE))
```

#checking Na values. There are not!
```{r,echo=TRUE}
mp <- aggr(df_new, col=c('navyblue','yellow'),
                   numbers=TRUE, sortVars=TRUE,
                   labels=names(df_new), cex.axis=.7,
                   gap=3, ylab=c("Missing data","Pattern"))
s<-summary(mp)

is.null(df_new)
```


#outcome variable (target)
```{r,echo=TRUE}
unique(df_new$Outcome)

df_new$y<-factor(df_new$Outcome, levels=c("Sa mat.stetom", "Sa povredjenim", "Sa poginulim"), ordered=TRUE)

levels(df_new$y)[1]<-"Material_damage"
levels(df_new$y)[2]<-"Injury"
levels(df_new$y)[3]<-"Death"

table(df_new$y) 
```

#A finally data set
```{r,echo=TRUE}
df_f<-df_new[,-c(1:2,10:13)]

for (i in c(2,4,6:7)){
  df_f[,i] <- as.numeric(as.character(df_f[,i]))
}

df_f$Month<-factor(df_f$Month, levels=c("January", "February"), ordered=TRUE)
df_f$month<-as.numeric(df_f$Month)

df_f$Date_Time<-with(df_f, ymd_h(paste(Year, Month, Day, Hour, sep= ' ')))

str(df_f)
head(df_f, 10)
```

Let's see the frequancy of accidents per months in year 2019
According to results there is no difference between frequencies of accidents.
```{r,echo=TRUE}
df_f_month<-df_f %>%
  group_by(Month, y) %>%
  summarise(n = n()) %>%
  mutate(Percent = round(n/sum(n)*100,2))

head(df_f_month,10)

m <- df_f_month%>%
      dplyr::select(Month, y, n)%>%
      spread(Month, n)

m <- m%>%
      as.data.frame()

Xsq <- chisq.test(m[-1], simulate.p.value = T)

m$p_value<-Xsq$p.value
m %>% 
  kable(booktabs = T,longtable = TRUE, caption = "Accidents during January and February") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
  collapse_rows(columns = 4)

ggplot(data = df_f_month, 
       aes(x = Month, y = Percent, color = y, fill = y)) +
  geom_bar(stat = "identity", position = "dodge", width = .5) + 
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  theme_bw()
```

Let's see the frequancy of accidents per months and days in year 2019
```{r,echo=TRUE}
df_f_January_day<-df_f %>%
  filter(Month=="January") %>%
  group_by(Day, y) %>%
  summarise(n = n()) %>%
  mutate(Percent = round(n/sum(n)*100,2))

#head(df_f_January_day,10)

ggplot(data = df_f_January_day, 
       aes(x = Day, y = Percent, color = y, fill = y)) +
  geom_bar(stat = "identity", position = "dodge", width = .5) + 
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  ggtitle("January")+
  theme_bw()

df_f_February_day<-df_f %>%
  filter(Month=="February") %>%
  group_by(Day, y) %>%
  summarise(n = n()) %>%
  mutate(Percent = round(n/sum(n)*100,2))

#head(df_f_February_day,10)

ggplot(data = df_f_February_day, 
       aes(x = Day, y = Percent, color = y, fill = y)) +
  geom_bar(stat = "identity", position = "dodge", width = .5) + 
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  ggtitle("February")+
  theme_bw()
```

Now, we will make two periods per months, I period: 1-15., II period: 16-31.
#Period I
No differences!
```{r,echo=TRUE}
df_f$period<-factor(ifelse(df_f$Day <= 15, 1, 2))

df_f_month_period1<-df_f %>%
  filter(period==1) %>%
  group_by(Month, y) %>%
  summarise(n = n()) %>%
  mutate(Percent = round(n/sum(n)*100,2))

m <- df_f_month_period1%>%
      dplyr::select(Month, y, n)%>%
      spread(Month, n)

m <- m%>%
      as.data.frame()

Xsq <- chisq.test(m[-1], simulate.p.value = T)

m$p_value<-Xsq$p.value
m %>% 
  kable(booktabs = T,longtable = TRUE, caption = "Accidents during January and February from 1st until 15th") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
  collapse_rows(columns = 4)

ggplot(data = df_f_month_period1, 
       aes(x = Month, y = Percent, color = y, fill = y)) +
  geom_bar(stat = "identity", position = "dodge", width = .5) + 
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  ggtitle("Period I")+
  theme_bw()
```

#Period II
No differences!
```{r,echo=TRUE}
df_f_month_period2<-df_f %>%
  filter(period==2) %>%
  group_by(Month, y) %>%
  summarise(n = n()) %>%
  mutate(Percent = round(n/sum(n)*100,2))

m <- df_f_month_period2%>%
      dplyr::select(Month, y, n)%>%
      spread(Month, n)

m <- m%>%
      as.data.frame()
m[is.na(m)] <- 0

Xsq <- chisq.test(m[-1], simulate.p.value = T)

m$p_value<-Xsq$p.value
m %>% 
  kable(booktabs = T,longtable = TRUE, caption = "Accidents during January and February from 11th until 20th") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
  collapse_rows(columns = 4)

ggplot(data = df_f_month_period2, 
       aes(x = Month, y = Percent, color = y, fill = y)) +
  geom_bar(stat = "identity", position = "dodge", width = .5) + 
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  ggtitle("Period II")+
  theme_bw()
```




Make data more interesant! Import weather data for Belgrade 01.01.2019-28.02.2019.

```{r,echo=TRUE}
 bgdWeather2019 <- read.csv('C:\\Users\\Vera\\Desktop\\history_export_2019-08-05T10_18_42.csv', na="", sep=";")
colnames(bgdWeather2019) <- gsub("<br />","", colnames(bgdWeather2019), fixed = T)
#str(bgdWeather2019)
bgdWeather2019$Date_Time<-with(bgdWeather2019, ymd_h(paste(Year, Month, Day, Hour, sep= ' ')))
 
df_big <- left_join(df_f, bgdWeather2019[,c(6:10)],
                      by = c("Date_Time" = "Date_Time"))
str(df_big)
is.null(df_big)
```


Do we have overdispersed or underdispersed count data. Over and choose negative binomial model.
```{r, echo=TRUE}
with(df_f_month, tapply(n, y, function(x) {
    sprintf("M (SD) = %1.2f (%1.2f)", mean(x, na.rm = T), var(x,na.rm = T))
}))
ggplot(df_f_period1, aes(n, fill = y)) + geom_histogram(binwidth = 1) + facet_grid(y ~ 
    ., margins = TRUE, scales = "free")
```

implementing the model
```{r, echo=TRUE}
df_f_month1<-df_big %>%
  group_by(Month, y) %>%
  summarise(Temperature = mean(Temperature...2.m.above.gnd.),n = n()) %>%
  mutate(Percent = round(n/sum(n)*100,2))
df_f_month1$y<-as.character(df_f_month1$y)
fit3 <- glm.nb(n ~ y+Temperature, data = df_f_month1, link = log)
coef(fit3, matrix = TRUE)
summary(fit3)
```
The indicator variable shown as yInjury is the expected difference in log count between group Injury and the reference group (Death). The expected log count for level Injury is 4.32 higer than the expected log count for level Death.
The indicator variable shown as yMaterial_damage is the expected difference in log count between group Material_damage and the reference group (Death). The expected log count for level Material_damage is 5.7 higer than the expected log count for level Death. 
The variable Temperature has a coefficient of -0.03, which is statistically significant. This means that for each one-unit dencrease in Temperature, the expected log count of the number of accidents increases by 0.03. 

#Prediction
First, we can look at predicted counts for each value of target variable while holding wind speed at its mean.
```{r, echo=TRUE}
newdf <- data.frame(Temperature = mean(df_f_month1$Temperature), y = factor(1:3, levels = 1:3, 
    labels = levels(as.factor(df_f_month1$y))))
newdf$yhat <- predict(fit3, newdf, type = "response")
newdf
```



From the results above, we see that the predicted number of accident Death is about 3.6, holding wind speed at its mean. The predicted number of accident Injury is about 268, and the predicted number of accident Material demage is about 1059.

Let's see how prediction looks at entire range temperature values.

```{r, echo=TRUE}
newdf2 <- data.frame(
  Temperature = rep(seq(from = min(df_f_month1$Temperature), to = max(df_f_month1$Temperature), length.out = 100), 3),
  y = factor(rep(1:3, each = 100), levels = 1:3, labels =
  levels(as.factor(df_f_month1$y))))

newdf2 <- cbind(newdf2, predict(fit3, newdf2, type = "link", se.fit=TRUE))
newdf2 <- within(newdf2, {
  y_pred <- exp(fit)
  LL <- exp(fit - 1.96 * se.fit)
  UL <- exp(fit + 1.96 * se.fit)
})

ggplot(newdf2, aes(Temperature,y_pred)) +
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = y), alpha = .25) +
  geom_line(aes(colour = y), size = 2) +
  labs(x = "Temperature", y = "Predicted outcome")
```